An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-10-03 paper/paper1/ms.tex
remove all the edit flags, switch to twocolumn

\documentclass[trackchanges]{aastex631}
\usepackage{hyperref}
\bibliographystyle{aasjournal}
%\turnoffedit
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{censor}
\usepackage[arrowdel]{physics}
\usepackage{mathtools}
\usepackage{outlines}
\usepackage{lipsum}  
\usepackage{amsmath,bm}

@@

A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.  Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} \edit1{and $\partial$\texttt{lux} \citep{10.1117/12.2629774}}, \edit1{stellar surface modeling with \texttt{starry} \citep{2021AJ....162..123L}}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.  Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.  \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.  The \texttt{ExoJAX} framework has only $\sim$100 tunable parameters that describe the fundamental physical properties controlling a \edit2{stratified} brown dwarf atmosphere.  These two autodiff-aware frameworks span the extreme ends of non-parametric and parametric modeling for spectroscopy.

@@

\edit1{Our first stage is to \emph{clone} a precomputed synthetic spectrum.  Cloning means to mimic, emulate, or otherwise approximate the appearance of a discretely sampled spectrum with some function or combination of functions.  The subject of the cloning shall be a noise-free spectrum generated from physics-based atmosphere models.  The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.}

@@

We place mild demands on the precomputed spectra. They should have sporadic regions of discernable continuum \edit2{(or pseudocontinuum)} devoid of lines, and the continuum shape should vary smoothly in wavelength. The spectral lines or pseudo-lines should be resolved (and not sub-sampled). Many precomputed synthetic stellar spectra will meet these criteria. \edit2{For those spectra that do not meet these criteria, the method should still work with some additional fine-tuning of the pre-processing steps that follow.  M dwarfs in the red-optical, solar-like stars in the blue visible, and brown dwarfs at virtually any wavelength may fall into this category.}

We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.  \edit1{The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out of- the region of interest.}  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars. A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.

@@

\noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.  \edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar output) or matrix multiplication (matrix output).  This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}

@@

Next, we identify the spectral lines. We apply the \edit1{\texttt{find\_peaks} local-maximum-finding algorithm implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} on the negative of the spectrum, in order to find the local minima}.  We select a ``topographic prominence threshold'' $P_{\rm rom} \in (0.005, 0.02)$, \edit1{defined as the vertical distance between the peak and the local baseline.  The baseline is defined with topographic heuristics that act as effective pseudo-continuum-detectors.}  This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines. The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved band heads provided that the individual lines exceed the prominence threshold in their local region. The number of lines $N_{\mathrm{lines}}$ depends on the bandwidth, prominence, and the intrinsic properties of the input spectrum, principally effective temperature and metallicity.

@@

\subsection{Goodness of fit metric and gradients}
The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths, lineshapes, and amplitudes. Instead, we tune the parameters of the model, starting from these coarse initial values. This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.  \edit2{The threshold-based line-finding algorithm performs very well at finding the line center positions, so in practice, we can fix the line center wavelengths.}  We, therefore, default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary. We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:

@@

As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the clone model has over 21,000 free parameters. Fitting more than about 300 parameters is difficult with conventional optimizers\edit1{, and so historically the practice of finding the best settings for all 21,000 parameters would have been considered hopeless, a fool's errand.  A new category of ``gradient-based'' optimizers offers a breakthrough.  Gradient-based optimizers have virtually no restriction on the number of parameters, scaling to functions of possibly millions or billions of tuning parameters.  Gradient-based optimizers get vastly more information per likelihood evaluation: $N_\mathrm{params}+1$ pieces of information-- the scalar value of the loss evaluated for some setting, \emph{and} the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.}  The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.

\edit1{Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit. This compass needle resides in, say, 21,000-dimensional space, instead of our mere 3 dimensions of physical space.  Conventional gradient-free optimizers evaluate the Jacobian through approximate finite differences, which result in numerical instabilities.  Those instabilities accrue and break down at around the 300-dimension mark.  So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.}

\edit1{We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.  In Full-Batch Gradient Descent, the entire dataset is evaluated in Equation \ref{simpleLikelihood}, as opposed to only a portion of the dataset in the Stochastic counterpart.  Appendix \ref{minibatches} discusses the tradeoffs and rationale for ``Full-Batch''.}

@@

We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center. Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra. The choice of $N_{\rm cut}$ is nuanced.  It should be set large enough that truncation effects are not seen for the broadest lines.  But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting. Extreme Doppler shifts could hypothetically send line \edit1{centers} entirely outside the extents of $N_{\rm cut}$ if set too low.  We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA. We coerce all wing cuts to be the same number of pixels, typically 6000 pixels, $\sim30-60~$\AA~ for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line. We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $6000$ pixels per line and assume zeros everywhere else.

@@

Second, a conspicuous flaw occurs in the line cores of relatively narrow lines.  The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.  \edit1{The residuals, therefore, exhibit a \texttt{W}-shape ringing artifact.  The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.  Recall that the cloning occurs at native spectral resolution, much greater than typical instrumental resolution.  The \texttt{W}-shape ringing residuals will tend to cancel each other out once the clone model undergoes convolution with an instrumental profile (Section \ref{jointModel}).  We ascribe this flaw to the adoption of approximate pseudo-Voigt profiles rather than exact Voigt profiles.  High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.}

\edit1{The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.  Even exact Voigt profiles inadequately represent true atmospheric opacities. Typical stellar and substellar atmosphere models compute opacities for a stratified atmosphere consisting of tens or hundreds of locally isothermal, isobaric atmospheric layers.  Each of those layers may exhibit a local per-line opacity that matches a Voigt profile.  But the cumulative product of many slightly-different Voigt profiles is not exactly equal to any single Voigt profile.  So the \emph{blas\'e} approach of asserting lines as Voigt profiles can be best understood as a phenomenological model-- that lines tend to look Voigt-like, even if they are non-Voigt in detail.  In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.  Such an approach is not yet implemented and may be an enticing avenue for future research.}

Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen lines and neutral alkali metal lines.  The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile, even an exact one.  Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.  \edit1{Nearby} narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike. This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.

@@

Equation \ref{eqn_final_model} has $\sim21,000$ tunable parameters from the star, $\sim9,000$ tunable parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.  That adds up to about $30,002$ model parameters.  The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration: $30,003$.  \edit2{The pseudocontinuum polynomial term $\mathsf{P}$ can also be tuned to refine the continuum placement initialized at the pre-processing stage, yielding an additional 5-15 parameters, depending on the choice of polynomial order.}

@@

\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$  \emph{blas\'e} line represents its (neighbor-weighted) Equivalent Width (EW).  That means that  \emph{blas\'e} can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.  Even more, the catalog would have some resilience to rotational broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.  Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.   \emph{Blas\'e} does not have these limitations.  The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained conventionally.  This approach may suit stellar abundance applications, such as Galactic archaeology and \edit1{chemically peculiar stars \citep[\emph{e.g. HgMn stars,}][]{2020MNRAS.496..832C}}.

@@

\edit1{\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.  \emph{Blas\'e} lossily recreates a spectrum, while having to store less information.  In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.}

\emph{``Surrogate modeling / Simulation Intelligence''}-- \edit1{Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.  The rapidly evolving field of machine learning may therefore classify the approach here using a few related terms.  The phrases surrogate modeling, emulation, discrepancy modeling, transfer learning, simulation intelligence, and physics-informed machine learning could describe some aspects of \emph{blas\'e}, depending on the ultimate application by the end-user.  The surrogate modeling approach has been applied with mixed success for M-dwarfs \citep{2020A&A...642A..22P} and brown dwarfs \citep{2020PASP..132d4502J}.  The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}. The key idea is that information from expensive physics-based simulations can be distilled by machine learning and then applied cheaply to new examples.  The performance of these surrogates should improve when more scientific domain knowledge is included at training time \citep{2018ApJ...869L...7A}.}

\emph{``Narrowing the Synthetic Gap'}-- \edit2{Our approach to pretraining on synthetic spectra and comparing to data resembles the technique of \citet{2020A&A...642A..22P}. There, a convolutional neural network is pretrained on PHOENIX spectra, and then applied to CARMENES spectra of M dwarfs.  \citet{2020A&A...642A..22P} designate the difference in appearance between the synthetic spectra and observed spectra as the ``synthetic gap'', showing that this gap leads to inaccurate metallicity inferences.  \emph{Blas\'e} shows a path towards shrinking the synthetic gap, with the prospect of re-training convolutional neural networks with semi-empirical spectra produced via the method described here, assuming high-fidelity metallicity labels can be obtained through co-natal binary studies.}

@@

\edit1{One key limitation stems from the need for autodiff to ``know about'' or sense each step of the end-to-end computation.  Practically speaking, this requirement means \texttt{blase} had to be written in a single differentiable (\emph{i.e.} autodiff-aware) programming framework.  This requirement is so strong as to forbid the intermingling of frameworks and languages.  The choice of which machine learning framework to adopt must therefore be made up-front, with some risk of \emph{vendor lock-in} once that choice has been set.  We considered three such differentiable frameworks: \texttt{TensorFlow}, \texttt{JAX}, and \texttt{PyTorch}.  We chose \texttt{PyTorch} because of its ease of use and its ability to structure differentiable models in an object-oriented style familiar within the \texttt{Python} community, with robust optimization support.  PyTorch had already been applied to several astronomy projects by this time, including \texttt{Exonet} \citep{2018ApJ...869L...7A}, \texttt{PySR} \citep{pysr}, and \texttt{MPoL} \citep{mpol}.  Ultimately, the choice of \texttt{PyTorch} made it easy to employ Sparse Matrices.}

\edit1{This vendor lock-in gives rise to a frustrating incompatibility among exoplanet-relevant autodiff applications.  For example, no differentiable piece of code in \texttt{blase} can call an \texttt{exoplanet} orbital radial velocity, a \texttt{starry} surface map, an \texttt{astropy} spectral coordinate transformation, a \texttt{wobble} spectral residual model, or an \texttt{ExoJAX} Voigt profile.  All of those tools use either \texttt{Theano}, \texttt{TensorFlow}, \texttt{Numpy}, or \texttt{JAX}, which are incompatible with \texttt{PyTorch}.  We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has been entered.  Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.}

\edit1{It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.  In practice, such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.  Each framework offers some strengths and demerits, and exploring those tradeoffs will continue to yield new experimental approaches.  For example, \texttt{JAX} may offer speedups attributable to Just In Time (JIT) compilation, at the expense of a more rigid functional programming style.  Access to scalable Gaussian Process implementations may be another deciding factor \citep{2022arXiv220908940A}.}

@@

\noindent \edit1{Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.  So instead, an educated guess for $\lambda_\mathrm{c,m}$ could be made based on the data centroid.  }

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{hyperref}
\bibliographystyle{aasjournal}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{censor}
\usepackage[arrowdel]{physics}
\usepackage{mathtools}
\usepackage{outlines}
\usepackage{lipsum}  
\usepackage{amsmath,bm}

@@

A key new enabling technology breaks these classical tradeoffs in data-model comparisons for astronomical spectroscopy.  Automatic differentiation \citep[``autodiff'' or ``autograd'',][]{2015arXiv150205767G, 2016PhDT.......317M} and its affiliated backpropagation algorithm \citep{kelley1960,Linnainmaa1976TaylorEO, 1986Natur.323..533R, 1990JGCD...13..926D} has revolutionized machine learning and neural network architecture design, and is increasingly applied in astrophysical data analysis contexts, \emph{e.g.} kernel phase coronography with \texttt{morphine} \citep{2021ApJ...907...40P,2021JOSAB..38.2465W} and $\partial$\texttt{lux} \citep{10.1117/12.2629774}, stellar surface modeling with \texttt{starry} \citep{2021AJ....162..123L}, and exoplanet orbit fitting with \texttt{exoplanet} \citep{2021JOSS....6.3285F}.  Of the spectroscopy frameworks mentioned above, the \texttt{TensorFlow}-based \citep{tensorflow2015-whitepaper} \texttt{wobble} and the \texttt{JAX}-based \citep{jax2018github} \texttt{ExoJAX} employ autodiff technology.  \texttt{wobble} treats each pixel as a tunable control point, producing $\sim10^5$ parameters for a modern stellar spectrum.  The \texttt{ExoJAX} framework has only $\sim$100 tunable parameters that describe the fundamental physical properties controlling a stratified brown dwarf atmosphere.  These two autodiff-aware frameworks span the extreme ends of non-parametric and parametric modeling for spectroscopy.

@@

Our first stage is to \emph{clone} a precomputed synthetic spectrum.  Cloning means to mimic, emulate, or otherwise approximate the appearance of a discretely sampled spectrum with some function or combination of functions.  The subject of the cloning shall be a noise-free spectrum generated from physics-based atmosphere models.  The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.

@@

We place mild demands on the precomputed spectra. They should have sporadic regions of discernable continuum (or pseudocontinuum) devoid of lines, and the continuum shape should vary smoothly in wavelength. The spectral lines or pseudo-lines should be resolved (and not sub-sampled). Many precomputed synthetic stellar spectra will meet these criteria. For those spectra that do not meet these criteria, the method should still work with some additional fine-tuning of the pre-processing steps that follow.  M dwarfs in the red-optical, solar-like stars in the blue visible, and brown dwarfs at virtually any wavelength may fall into this category.

We truncate the red and blue limits of the precomputed synthetic spectrum to match a high-bandwidth echelle spectrograph, extended with a buffer at the edges of size $\pm \Delta \lambda_{\mathrm{buffer}}$.  The buffers are needed for two reasons: 1) to include the influence of spectral lines whose centers lie just outside the region-of-interest, but whose wings are broad enough to affect the neighboring pixels; and 2) to account for astrophysical radial velocity shifts that may send line centers into- and out of- the region of interest.  We set the value of the buffers to account for plausible radial velocity and rotational broadening of real stars. A generous buffer of $v \sin{i} < 500 \;\kmps$ and $|RV|<500 \;\kmps$ yields a typical buffer of about 30~\AA.

@@

\noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.  For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar output) or matrix multiplication (matrix output).  This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}

@@

Next, we identify the spectral lines. We apply the \texttt{find\_peaks} local-maximum-finding algorithm implemented in \texttt{SciPy} \citep{2020SciPy-NMeth} on the negative of the spectrum, in order to find the local minima.  We select a ``topographic prominence threshold'' $P_{\rm rom} \in (0.005, 0.02)$, defined as the vertical distance between the peak and the local baseline.  The baseline is defined with topographic heuristics that act as effective pseudo-continuum-detectors.  This threshold dictates the number of lines that will be modeled: a lower prominence finds more, weaker lines, and a larger prominence finds fewer, deeper lines. The prominence algorithm successfully finds lines that reside on top of broad line wings, or unresolved band heads provided that the individual lines exceed the prominence threshold in their local region. The number of lines $N_{\mathrm{lines}}$ depends on the bandwidth, prominence, and the intrinsic properties of the input spectrum, principally effective temperature and metallicity.

@@

\subsection{Goodness of fit metric and gradients}
The model evaluated with its coarse initial values would have terrible performance: it would only vaguely resemble the synthetic spectral model, with up to $\pm 50\%$ undulations from the inexact assignment of widths, lineshapes, and amplitudes. Instead, we tune the parameters of the model, starting from these coarse initial values. This model has between $N_{\mathrm{lines}}\times 1$ and $N_{\mathrm{lines}}\times 4$ free parameters, depending on how many of the 4 physical line properties you wish to tune.  The threshold-based line-finding algorithm performs very well at finding the line center positions, so in practice, we can fix the line center wavelengths.  We, therefore, default to fitting three parameters per line, where the center wavelength is held fixed and the amplitude, width, and lineshape are allowed to vary. We minimize a scalar ``goodness-of-fit'' metric, \emph{aka} loss scalar $\mathcal{L}$, chosen as the mean squared error (MSE), which is proportional to $\chi^2$, the sum of the squares of the residual vector $\mathsf{R} \equiv \mathsf{S}-\mathsf{S}_{\rm clone}$ but has no notion of per-pixel noise since the precomputed synthetic spectrum has no uncertainty:

@@

As seen in Figure \ref{fig_Nlines_vs_teff}, the number of lines can exceed 7,000, meaning the clone model has over 21,000 free parameters. Fitting more than about 300 parameters is difficult with conventional optimizers, and so historically the practice of finding the best settings for all 21,000 parameters would have been considered hopeless, a fool's errand.  A new category of ``gradient-based'' optimizers offers a breakthrough.  Gradient-based optimizers have virtually no restriction on the number of parameters, scaling to functions of possibly millions or billions of tuning parameters.  Gradient-based optimizers get vastly more information per likelihood evaluation: $N_\mathrm{params}+1$ pieces of information-- the scalar value of the loss evaluated for some setting, \emph{and} the derivative of the loss scalar with respect to each of the parameters, the so-called \emph{Jacobian}: $\grad \mathcal{L} = (\frac{\partial \mathcal{L}}{\partial a_j}, \frac{\partial \mathcal{L}}{\partial \sigma_j}, \cdots, \frac{\partial \mathcal{L}}{\partial \gamma_j})$.  The Jacobian indicates how the MSE would decrease with a change in the parameter-of-interest, or put simply ``which way and by how much'' you have to change each individual line property to get a better fit.

Geometrically, the $\grad \mathcal{L}$ Jacobian resembles a compass needle guiding the direction to the best fit. This compass needle resides in, say, 21,000-dimensional space, instead of our mere 3 dimensions of physical space.  Conventional gradient-free optimizers evaluate the Jacobian through approximate finite differences, which result in numerical instabilities.  Those instabilities accrue and break down at around the 300-dimension mark.  So non-gradient-based optimization can be pictured as having only a coarse needle---assembled from approximate finite differences---that begins to spin aimlessly after too many dimensions.

We adopt ``Full-batch'' Gradient Descent optimization, as opposed to the more popular Stochastic Gradient Descent \citep[SGD,][]{2016arXiv160904747R}.  In Full-Batch Gradient Descent, the entire dataset is evaluated in Equation \ref{simpleLikelihood}, as opposed to only a portion of the dataset in the Stochastic counterpart.  Appendix \ref{minibatches} discusses the tradeoffs and rationale for ``Full-Batch''.

@@

We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center. Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra. The choice of $N_{\rm cut}$ is nuanced.  It should be set large enough that truncation effects are not seen for the broadest lines.  But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting. Extreme Doppler shifts could hypothetically send line centers entirely outside the extents of $N_{\rm cut}$ if set too low.  We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA. We coerce all wing cuts to be the same number of pixels, typically 6000 pixels, $\sim30-60~$\AA~ for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line. We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $6000$ pixels per line and assume zeros everywhere else.

@@

Second, a conspicuous flaw occurs in the line cores of relatively narrow lines.  The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.  The residuals, therefore, exhibit a \texttt{W}-shape ringing artifact.  The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.  Recall that the cloning occurs at native spectral resolution, much greater than typical instrumental resolution.  The \texttt{W}-shape ringing residuals will tend to cancel each other out once the clone model undergoes convolution with an instrumental profile (Section \ref{jointModel}).  We ascribe this flaw to the adoption of approximate pseudo-Voigt profiles rather than exact Voigt profiles.  High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.

The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.  Even exact Voigt profiles inadequately represent true atmospheric opacities. Typical stellar and substellar atmosphere models compute opacities for a stratified atmosphere consisting of tens or hundreds of locally isothermal, isobaric atmospheric layers.  Each of those layers may exhibit a local per-line opacity that matches a Voigt profile.  But the cumulative product of many slightly-different Voigt profiles is not exactly equal to any single Voigt profile.  So the \emph{blas\'e} approach of asserting lines as Voigt profiles can be best understood as a phenomenological model-- that lines tend to look Voigt-like, even if they are non-Voigt in detail.  In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.  Such an approach is not yet implemented and may be an enticing avenue for future research.

Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen lines and neutral alkali metal lines.  The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile, even an exact one.  Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.  Nearby narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike. This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.

@@

Equation \ref{eqn_final_model} has $\sim21,000$ tunable parameters from the star, $\sim9,000$ tunable parameters from the Earth's atmosphere, plus $v\sin{i}$ and $RV$.  That adds up to about $30,002$ model parameters.  The resolving power may also be treated as tunable if it is not known or varies slightly with \emph{e.g.} seeing, slit-or-fiber illumination, or instrumental configuration: $30,003$.  The pseudocontinuum polynomial term $\mathsf{P}$ can also be tuned to refine the continuum placement initialized at the pre-processing stage, yielding an additional 5-15 parameters, depending on the choice of polynomial order.

@@

\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$  \emph{blas\'e} line represents its (neighbor-weighted) Equivalent Width (EW).  That means that  \emph{blas\'e} can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.  Even more, the catalog would have some resilience to rotational broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.  Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.   \emph{Blas\'e} does not have these limitations.  The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained conventionally.  This approach may suit stellar abundance applications, such as Galactic archaeology and chemically peculiar stars \citep[\emph{e.g. HgMn stars,}][]{2020MNRAS.496..832C}.

@@

\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.  \emph{Blas\'e} lossily recreates a spectrum, while having to store less information.  In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.

\emph{``Surrogate modeling / Simulation Intelligence''}-- Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.  The rapidly evolving field of machine learning may therefore classify the approach here using a few related terms.  The phrases surrogate modeling, emulation, discrepancy modeling, transfer learning, simulation intelligence, and physics-informed machine learning could describe some aspects of \emph{blas\'e}, depending on the ultimate application by the end-user.  The surrogate modeling approach has been applied with mixed success for M-dwarfs \citep{2020A&A...642A..22P} and brown dwarfs \citep{2020PASP..132d4502J}.  The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}. The key idea is that information from expensive physics-based simulations can be distilled by machine learning and then applied cheaply to new examples.  The performance of these surrogates should improve when more scientific domain knowledge is included at training time \citep{2018ApJ...869L...7A}.

\emph{``Narrowing the Synthetic Gap'}-- Our approach to pretraining on synthetic spectra and comparing to data resembles the technique of \citet{2020A&A...642A..22P}. There, a convolutional neural network is pretrained on PHOENIX spectra, and then applied to CARMENES spectra of M dwarfs.  \citet{2020A&A...642A..22P} designate the difference in appearance between the synthetic spectra and observed spectra as the ``synthetic gap'', showing that this gap leads to inaccurate metallicity inferences.  \emph{Blas\'e} shows a path towards shrinking the synthetic gap, with the prospect of re-training convolutional neural networks with semi-empirical spectra produced via the method described here, assuming high-fidelity metallicity labels can be obtained through co-natal binary studies.

@@

One key limitation stems from the need for autodiff to ``know about'' or sense each step of the end-to-end computation.  Practically speaking, this requirement means \texttt{blase} had to be written in a single differentiable (\emph{i.e.} autodiff-aware) programming framework.  This requirement is so strong as to forbid the intermingling of frameworks and languages.  The choice of which machine learning framework to adopt must therefore be made up-front, with some risk of \emph{vendor lock-in} once that choice has been set.  We considered three such differentiable frameworks: \texttt{TensorFlow}, \texttt{JAX}, and \texttt{PyTorch}.  We chose \texttt{PyTorch} because of its ease of use and its ability to structure differentiable models in an object-oriented style familiar within the \texttt{Python} community, with robust optimization support.  PyTorch had already been applied to several astronomy projects by this time, including \texttt{Exonet} \citep{2018ApJ...869L...7A}, \texttt{PySR} \citep{pysr}, and \texttt{MPoL} \citep{mpol}.  Ultimately, the choice of \texttt{PyTorch} made it easy to employ Sparse Matrices.

This vendor lock-in gives rise to a frustrating incompatibility among exoplanet-relevant autodiff applications.  For example, no differentiable piece of code in \texttt{blase} can call an \texttt{exoplanet} orbital radial velocity, a \texttt{starry} surface map, an \texttt{astropy} spectral coordinate transformation, a \texttt{wobble} spectral residual model, or an \texttt{ExoJAX} Voigt profile.  All of those tools use either \texttt{Theano}, \texttt{TensorFlow}, \texttt{Numpy}, or \texttt{JAX}, which are incompatible with \texttt{PyTorch}.  We use \texttt{SciPy}, but only for preprocessing, before the differentiable model has been entered.  Some heuristics exist for interchanging code among the frameworks, but they pose some implementation complexity and/or some performance hit.

It may appear attractive to coordinate the relatively small autodiff-for-astrophysics community to adopt a single autodiff-able machine learning framework to facilitate compatibility.  In practice, such coordination may be ill-fated, since the rapid pace of machine learning development has not yet settled on a single best way forward.  Each framework offers some strengths and demerits, and exploring those tradeoffs will continue to yield new experimental approaches.  For example, \texttt{JAX} may offer speedups attributable to Just In Time (JIT) compilation, at the expense of a more rigid functional programming style.  Access to scalable Gaussian Process implementations may be another deciding factor \citep{2022arXiv220908940A}.

@@

\noindent Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.  So instead, an educated guess for $\lambda_\mathrm{c,m}$ could be made based on the data centroid.

In either case, the initialization of $A_m, \sigma_m, \gamma_m$ would be a matter of taste, and they would therefore be difficult to regularize, at least in the manner that we have described so far.  For these reasons, an even better strategy may be to treat the data residuals not as lines, but in some other basis.  For example, it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.  These questions remain an open research area.
